On the formation of H$\alpha$ line emission around classical T Tauri stars

### 2006-01-26 TTau_Halpha/kurosawa.tex
Major structual change. Now more or less complte. The discussion sections may need  revisions. Some plots need to be replaced after the model curretly running finishes.  (RK)
git-svn-id: https://repository.astro.ex.ac.uk/torus/trunk@632 25464f21-6099-402c-b30a-640f1d40d38b

%% LyX 1.3 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[english,useAMS, usenatbib]{mn2e}
\usepackage[T1]{fontenc}
\usepackage[latin1]{inputenc}
\usepackage{times}
\usepackage{graphicx}
\usepackage{amssymb}

@@

  \caption{Basic model configuration of the disc-wind, magnetosphere
    hybrid model. The system consist of four components: (1)~the
    continuum source located at the origin $\left( o \right)$ of the
    cartesian coordinates $\left(x,y,z\right)$ -- the $y$-axis is into
    the paper, (2)~the magnetospheric accretion flow, (3)~the disc
    wind, and (4)~the accretion disc. The disc wind originates from
    the disc surface between $w_{i}=R_{\mathrm{wi}}$ and
    $w_{i}=R_{\mathrm{wo}}$ where $w_{i}$ is the distance from the $z$
    axis on the equatorial plane.  The wind source
    points ($S$), from which the stream lines diverges, are placed at
    distance $d$ above and below the star. The degree of wind
    collimation is controlled by changing the value of $d$.}

@@

We adopt the MA flow model of \citet{hartmann:1994},
as done  by \citet{muzerolle:2001} and by \citet*{symington:2005},
in which the gas accretion on to the stellar surface from the innermost
part of the accretion disc occurs through a dipolar stellar magnetic
field. The magnetic field is assumed to be so strong that the gas
flow does not affect the underlying magnetic field itself. As shown
in Fig.~\ref{ig:config_discwind} (region~2), the innermost radius ($R_{\mathrm{mi}}$)
of the magnetosphere at the equatorial plane ($z=0$) is assigned
to be same as the inner radius ($R_{\mathrm{di}}$) of the accretion
disc where the flow is truncated. In our models, $R_{\mathrm{mi}}$
and the outer radius ($R_{\mathrm{mo}}$) of the magnetosphere (at
the equatorial plane) are set to be $2.2\, R_{\sun}$ and $3.0\, R_{\sun}$
respectively. The geometry of the magnetic field/stream
lines is fixed for all calculations. We note that this magnetospheric geometry
is identical to the {}``small/wide'' model of \citet{muzerolle:2001}.

@@

For a point in the magnetosphere and the wind flows, the emissivity
and the opacity of the media are given as:
\begin{equation}
  \left\{ 
  \begin{array}{rcl}
    \eta_{\nu} & = & \eta_{\mathrm{c}}^{\mathrm{H}}+\eta_{l}^{\mathrm{H}}\\
    \chi_{\nu} & = & \chi_{c}^{\mathrm{H}}+\chi_{l}^{\mathrm{H}}
    +\sigma_{\mathrm{es}}
  \end{array}
  \right.
  \label{eq:emiss_opa_gas}
\end{equation}
where $\eta_{\mathrm{c}}^{\mathrm{H}}$ and $\eta_{l}^{\mathrm{H}}$ are
the continuum and line emissivity of
hydrogen. $\chi_{c}^{\mathrm{H}}$, $\chi_{l}^{\mathrm{H}}$, and
$\sigma_{\mathrm{es}}$ are the continuum, line opacity
(equation~\ref{eq:line_opacity}) of hydrogen, and the electron
scattering opacity. Similarly, for a point in the accretion disc,
\begin{equation}
  \left\{ 
  \begin{array}{rcl}
    \eta_{\nu} & = & 0\\
    \chi_{\nu} & = & \kappa_{\mathrm{abs}}^{\mathrm{dust}}
    +\kappa_{\mathrm{sca}}^{\mathrm{dust}}
  \end{array}
  \right.
  \label{eq:emiss_opa_dust}
\end{equation}
where $\kappa_{\mathrm{abs}}^{\mathrm{dust}}$ and
$\kappa_{\mathrm{sca}}^{\mathrm{dust}}$ are the dust absorption, and
scattering opacity which are calculated using the dust property
described in section~\ref{eq:grain-dist-function}.  We neglect the
emissivity from the disc: since the disc mass of CTTS are rather small
($\sim 0.01\,\mathrm{M_{\sun}}$) and low temperature ($<\sim
1600\,\mathrm{K}$), the continuum flux contribution at H$\alpha$
wavelength is expected to be negligible (e.g.~\citealt{chiang:1997}).

\section{Results}

\label{sec:Results}

\subsection{Magnetosphere only models}

\label{sub:model-acc}

@@

    \includegraphics[%
      clip,
      scale=0.45]{figures/disc_wind_acc_mdot_effect.eps}

  \caption{The effect of the mass-loss rate to mass-accretion rate
    ratio ($\mu=\dot{M}_{\mathrm{wind}}/\dot{M}_{\mathrm{acc}}$).  The
    H$\alpha$ profile computed with only the magnetospheric accretion
    (thick solid) is compared with the magnetosphere--wind--disc
    hybrid models for $\mu =$ 0.01 (solid), 0.025 (dash-dot), 0.1
    (dot) and 0.2 (dash). The magnetosphere used here has
    $T_{\mathrm{max}}=7500\,\Kelvin$ and
    $\dot{M}_{\mathrm{acc}}=10^{-7}\,\mathrm{M_{\sun}\,yr^{-1}}$.  The
    temperature and the acceleration parameter of the bipolar wind are
    $T_{\mathrm{wind}}=7000\,\Kelvin$ and $\beta=2.0$ respectively.
    All the models are computed with the inclination angle
    $i=55^{\circ}$. As $\mu$ increases the P-Cygni absorption deepens
    and the emission component becomes stronger. In general, the
    absorption is too strong for the models with $\mu>0.025$ compared
    to the observations (e.g., \citealt{reipurth:1996}).}

  \label{fig:effect_mdot_disc_wind_acc}

\subsection{Disc-wind only models}

\label{sub:result-discwind-knigge}

@@

  \caption{H$\alpha$ profiles computed with the disc-wind only.
     The wind mass-loss rate is fixed at
    $\dot{M}_{\mathrm{wind}}=10^{-8}\,\mathrm{M_{\sun}\,yr^{-1}}$, and the line
    profiles are computed with different combinations of the wind
    acceleration rate ($\beta$) and isothermal disc-wind temperature
    ($T_{\mathrm{wind}}$). The wind emission grows as the values of
    $\beta$ and $T_{\mathrm{wind}}$ increase.   As in the
    bipolar wind models (Figure~\ref{fig:atlas_wind_model}), the position
    of the wind absorption moves toward the line centre as $\beta$ becomes
    larger.     }

@@

@@

  \caption{The disc-wind, magnetosphere hybrid model. These models
  (solid lines) have the  same parameters as
    Figure~\ref{fig:atlas_discwind_only}, but  also
    include the magnetospheric accretion flow
    ($T_{\mathrm{\max}}=7500~\mathrm{K}$ and
    $\dot{M}_{\mathrm{acc}}=10^{-7}\,\mathrm{M_{\sun}\,yr^{-1}}$). While
    the emission from the magnetosphere dominates for the models with
    smaller $\beta$ and $T_{\mathrm{wind}}$, the emission from the wind dominates the
    profiles for models with larger $\beta$ and $T_{\mathrm{wind}}$.  The contribution
    of the magnetosphere can be judged from the profiles computed with
    the magnetosphere only (dotted lines).}

  \label{fig:atlas_discwind_acc_model}

@@

In this section, we examine profiles produced using the 
the disc wind outflow model described in
section~\ref{sub:model-disc-wind}. This type of outflow 
has been predicted by several MHD calculations (e.g. \citet{koenigl:2000};
 \citealt{krasnopolsky:2003}), and broadly consists of a wind with
its density increasing toward the equatorial plane.

GEOMETRY? REFERENCE MODEL?

The parameters used for the
central star are as in section~\ref{sub:Continuum-Source}, and the
disc-wind parameters are summarised in
Table~\ref{tab:std_discwind_parameters}.

Although the line is potentially
sensitive to the temperature structure of the wind, determination of a
self-consistent wind temperature is beyond the scope of this
paper. Readers are refereed to \citet{hartmann:1982} in which the wind
temperature structure is determined by balancing the radiative cooling
rate (assuming optically thin) with the MHD wave heating rate.
We pragmatically assume that the wind is isothermal at $T_{\mathrm{wind}}$.

Initially we examine the characteristics of the
H$\alpha$ profile  as a function of the wind acceleration parameter
$\beta$ and the isothermal wind temperature
$T_{\mathrm{wind}}$.  Figure~\ref{fig:atlas_discwind_only} shows profiles computed at  $i=55^{\circ}$
 for a range of wind temperatures
between $7000\,\mathrm{K}$ and $9000\,\mathrm{K}$, and  $\beta$ between $0.2$ and $2.0$.

Classical P~Cygni profiles are prominent
in the models with lower values of $T_{\mathrm{wind}}$ and $\beta$
(a colder and faster accelerating wind). As the value of $\beta$
increases, the position of the blue-shifted absorption component moves
toward the line centre -- this can be seen clearly in the
$T_{\mathrm{wind}}=7000\,\mathrm{K}$ models. The velocity position
 of the absorption component of the fastest accelerating wind
($\beta=0.2$) is almost at the terminal velocity ($v_{\infty}\sim
210\,\kmps$), that for the slowest wind acceleration model
($\beta=4.0$) is located close the line centre ($\sim 50\,\kmps$).

 As
one can see in Figure~\ref{fig:wind_vr_rho}, the column density of the
gas moving at $v_{\infty}$ increases as the wind acceleration becomes
larger since the gas moving at $v_{\infty}$ becomes closer to the
photosphere. At the location where the absorption occurs, the gas is
moving faster for the models with the faster wind acceleration models;
hence, the location of the absorption component is bluer as $\beta$
becomes smaller (as the wind acceleration faster). The figure also
shows that the intensity of the emission component increases as
$\beta$ increases for a given $T_{\mathrm{wind}}$. This is mainly
because the density near the photosphere increases as the value of
$\beta$ increases (Figure~\ref{fig:wind_vr_rho}); hence, the
emissivity of increases. Unlike the magnetospheric accretion only
models (Figure~\ref{fig:atlas_acc_models}), the inverse P~Cygni
profiles are not seen here, nor the blue-ward asymmetry. Because of
the absorption component in P~Cygni profile, most of the profiles show
red-ward asymmetry except for the low temperature wind models with
high inclinations (not shown here).

The effect of varying the wind mass-loss rate is shown in Figure~\ref{fig:effect_wind_mdot}.
The line is very sensitive to the mass-loss rate. With all other parameters
fixed constant, $\dot{M}_{\mathrm{wind}}$ acts as an scaling factor
of the wind density (equation~\ref{eq:wind_density_final}). The
important processes in forming H$\alpha$ emission are the
recombination process and the subsequent downward transition to $n=3$
level. Assuming the ionisation fraction remains constant, the
recombination rate is proportional to the square 
of wind density. We have found the line flux of the models in the
figure is approximately proportional to the square of the mass-loss
rate. This very sensitive nature of H$\alpha$ would make the line
a good candidate for a mass-loss rate indicator; however, without
solving the ambiguity of the emission component from the magnetospheric
accretion flow and the wind, and without knowing the detailed temperature
structure of the wind, it could provide us a misleading a mass-loss
rate estimate of a T Tauri star.

Figure~\ref{fig:atlas_discwind_only} shows the H$\alpha$ profiles
computed using only the disc-wind component
(c.f. Figure~\ref{fig:config_discwind}) for the same isothermal
disc-wind temperature ($T_{\mathrm{wind}}$) and the wind acceleration
parameter ($\beta$) ranges used for the bipolar wind models
(c.f. Figures~\ref{fig:atlas_wind_model} and
\ref{fig:atlas_disc_wind_acc_model}).

  The figure shows that the
wind emission increases as $\beta$ and $T_{\mathrm{wind}}$ increase. The
$\beta$ dependence arises simply because for large $\beta$ there is a
greater mass of wind material close to the star.  As with the bipolar
wind models, the location of the P-Cyg absorption component moves
toward the line centre as $\beta$ increases.

The morphology of the profile exhibited by the model changes from
FU-Ori type (c.f.~\citealt{reipurth:1996}) for the models with small
$\beta$ to the stronger emission types which are more commonly seen in
the observed H$\alpha$ (c.f.~\citealt{reipurth:1996}) as the value of
$\beta$ increases.  We find that the profile shapes are relatively
insensitive to the value of the wind collimation parameter $d$. The
dependency on the disc wind mass-loss rate of the profile shape is
similar to that of the bipolar wind models
(Figure~\ref{fig:effect_mdot_disc_wind_acc}).

\subsection{Disc--wind--magnetosphere  hybrid models}

\label{sub:result-discwind-hybrid}

We now consider the models computed with a combination of the
disc-wind and the magnetospheric accretion components. The
parameters used for the magnetosphere are set to be same as in the
case for Figure~\ref{fig:acc_inclination_effect}.  
The mass-loss rate of the disc wind is set to
$\dot{M}_{\mathrm{wind}}=10^{-8}\,\mathrm{M_{\sun}\, yr^{-1}}$
(i.e.~$\mu=\dot{M}_{\mathrm{wind}}/\dot{M}_{\mathrm{acc}}=0.1$).
Figure~\ref{fig:atlas_discwind_acc_model} shows the models profiles
computed using the same ranges of the disc-wind temperature and the
wind acceleration parameter as in the previous cases.

As in the disc-wind only models
(Figure~\ref{fig:atlas_discwind_only}), the location of the absorption
component moves toward the line centre as the value of $\beta$
increases.  The figure shows that in
the models with smaller $\beta$ and $T_{\mathrm{wind}}$, the line
emission from the magnetosphere dominates, while this situation reverses for the
models with larger $\beta$ and $T_{\mathrm{wind}}$. The model with
$\left(T_{\mathrm{wind}},\,\beta\right)=\left(9000\,\mathrm{K},\,2.0\right)$
results in an H$\alpha$ profile that is far too strong to be compatible
 with observations (\citet{reipurth:1996} and \citet{alencar:2000});
 however, the lower mass-loss rate models (while keeping $\mu$ constant)
 produce the profiles with the line strengths comparable to observations.
Although the models in Figure~\ref{fig:atlas_discwind_acc_model} are
computed with limited ranges of  $T_{\mathrm{wind}}$ and $\beta$, the
resulting profiles exhibit a wide variety of line profile shapes,
many of  which are similar to the types of H$\alpha$ profiles seen in the
observations (c.f.~\citealt{reipurth:1996}).

An examination of the profiles in Figures~\ref{fig:atlas_discwind_only} and
Figure~\ref{fig:atlas_discwind_acc_model} reveals some of problems
with uniqueness that must inevitably occur when trying to assign
an individual profile as wind or accretion dominated i.e. very similar profiles may arise
from very different circumstellar geometries. For example, the disc-wind
only model with $\left(T_{\mathrm{wind}},\,\beta\right)=\left(8000\,\mathrm{K},\,1.0\right)$
 (in Figures~\ref{fig:atlas_discwind_only}) and the  accretion dominated model with
 $\left(T_{\mathrm{wind}},\,\beta\right)=\left(9000\,\mathrm{K},\,0.5\right)$
(in Figure~\ref{fig:atlas_discwind_acc_model}) give very similar profiles.

The general characteristics of the profiles from the disc-wind, magnetosphere
hybrid model (Figure~\ref{fig:atlas_discwind_acc_model}) are very
similar to those of the disc, bipolar wind, magnetosphere hybrid
model (Figure~\ref{fig:atlas_disc_wind_acc_model}). The main difference
between these two sets of models is in the ratio of the mass-loss rate to the mass
accretion rate, $\mu$. While the models presented in this section
use $\mu=0.1$ which is consistent with the MHD jet models (e.g.~\citealt{koenigl:2000}),
the models in Figure~\ref{fig:atlas_disc_wind_acc_model} use much
smaller value ($\mu=0.01$).   The dependency of the model profiles on
the inclination angle for the two different wind models will be discussed
in the next section.

\begin{table}

\caption{A summary of the reference disc-wind model parameters used
in section~\ref{sub:result-discwind-knigge}. See also
section~\ref{sub:Alternative-model-disc-wind}.}

@@

\begin{tabular}{ccccccc}
\hline 
$d$&
$R_{\mathrm{wi}}$&
$R_{\mathrm{wo}}$&
$\dot{M}_{\mathrm{wind}}$&
$p$&
$f$&
$R_{\mathrm{s}}$\tabularnewline
$\left[R_{*}\right]$&
$\left[R_{*}\right]$&
$\left[\mathrm{au}\right]$&
$\left[M_{\sun}\,\mathrm{yr^{-1}}\right]$&
$[-]$&
$[-]$&
$\left[R_{*}\right]$\tabularnewline
\hline
$21.4$&
$3.0$&
$1.0$&
$10^{-8}$&
$-3/2$&
$2.0$&
$30$\tabularnewline
\hline
\end{tabular}

@@

\end{table}

@@

\subsection{H$\alpha$ as a mass accretion diagnostic}

>>>>

\documentclass[english,useAMS, usenatbib]{mn2e}
\usepackage[T1]{fontenc}
\usepackage[latin1]{inputenc}
\usepackage{times}
\usepackage{graphicx}
\usepackage{amssymb}

@@

  \caption{Basic model configuration of the disc-wind-magnetosphere
    hybrid model. The system consist of four components: (1)~the
    continuum source located at the origin $\left( o \right)$ of the
    cartesian coordinates $\left(x,y,z\right)$ -- the $y$-axis is into
    the paper, (2)~the magnetospheric accretion flow, (3)~the disc
    wind, and (4)~the accretion disc. The disc wind originates from
    the disc surface between $w_{i}=R_{\mathrm{wi}}$ and
    $w_{i}=R_{\mathrm{wo}}$ where $w_{i}$ is the distance from the $z$
    axis on the equatorial plane.  The wind source
    points ($S$), from which the stream lines diverges, are placed at
    distance $d$ above and below the star. The degree of wind
    collimation is controlled by changing the value of $d$.}

@@

We adopt the MA flow model of \citet{hartmann:1994},
as done  by \citet{muzerolle:2001} and by \citet*{symington:2005},
in which the gas accretion on to the stellar surface from the innermost
part of the accretion disc occurs through a dipolar stellar magnetic
field. The magnetic field is assumed to be so strong that the gas
flow does not affect the underlying magnetic field itself. As shown
in Fig.~\ref{fig:config_discwind} (region~2), the innermost radius ($R_{\mathrm{mi}}$)
of the magnetosphere at the equatorial plane ($z=0$) is assigned
to be same as the inner radius ($R_{\mathrm{di}}$) of the accretion
disc where the flow is truncated. In our models, $R_{\mathrm{mi}}$
and the outer radius ($R_{\mathrm{mo}}$) of the magnetosphere (at
the equatorial plane) are set to be $2.2\, R_{\sun}$ and $3.0\, R_{\sun}$
respectively. The geometry of the magnetic field/stream
lines is fixed for all calculations. We note that this magnetospheric geometry
is identical to the {}``small/wide'' model of \citet{muzerolle:2001}.

@@

For a point in the magnetosphere and the wind flows, the emissivity
and the opacity of the media are given as:
\begin{equation}
  \left\{ 
  \begin{array}{rcl}
    \eta_{\nu} & = & \eta_{\mathrm{c}}^{\mathrm{H}}+\eta_{l}^{\mathrm{H}}\\
    \chi_{\nu} & = & \chi_{c}^{\mathrm{H}}+\chi_{l}^{\mathrm{H}}
    +\sigma_{\mathrm{es}}
  \end{array}
  \right.
  \label{eq:emiss_opa_gas}
\end{equation}
where $\eta_{\mathrm{c}}^{\mathrm{H}}$ and $\eta_{l}^{\mathrm{H}}$ are
the continuum and line emissivity of
hydrogen. $\chi_{c}^{\mathrm{H}}$, $\chi_{l}^{\mathrm{H}}$, and
$\sigma_{\mathrm{es}}$ are the continuum, line opacity
(equation~\ref{eq:line_opacity}) of hydrogen, and the electron
scattering opacity. Similarly, for a point in the accretion disc,
\begin{equation}
  \left\{ 
  \begin{array}{rcl}
    \eta_{\nu} & = & 0\\
    \chi_{\nu} & = & \kappa_{\mathrm{abs}}^{\mathrm{dust}}
    +\kappa_{\mathrm{sca}}^{\mathrm{dust}}
  \end{array}
  \right.
  \label{eq:emiss_opa_dust}
\end{equation}
where $\kappa_{\mathrm{abs}}^{\mathrm{dust}}$ and
$\kappa_{\mathrm{sca}}^{\mathrm{dust}}$ are the dust absorption, and
scattering opacity which are calculated using the dust property
described in section~\ref{eq:grain-dist-function}.  We neglect the
emissivity from the disc: since the disc mass of CTTS are rather small
($\sim 0.01\,\mathrm{M_{\sun}}$) and low temperature ($\lesssim 
1600\,\mathrm{K}$), the continuum flux contribution at H$\alpha$
wavelength is expected to be negligible (e.g.~\citealt{chiang:1997}).

\section{Results}
\label{sec:Results}

\subsection{Magnetosphere only models}
\label{sub:model-acc}

@@

\subsection{Disc-wind only models}
\label{sub:result-discwind-knigge}

@@

  \caption{H$\alpha$ profiles computed with the disc-wind only.
     The wind mass-loss rate is fixed at
    $\dot{M}_{\mathrm{wind}}=10^{-8}\,\mathrm{M_{\sun}\,yr^{-1}}$, and the line
    profiles are computed with different combinations of the wind
    acceleration rate ($\beta$) and isothermal disc-wind temperature
    ($T_{\mathrm{wind}}$). The wind emission grows as the values of
    $\beta$ and $T_{\mathrm{wind}}$ increase.   The position
    of the wind absorption moves toward the line centre as $\beta$ becomes
    larger.   }

@@

In this section, we examine profiles produced using the 
the disc wind outflow model described in
section~\ref{sub:model-disc-wind}. 
The parameters used for the
central star are as in section~\ref{sub:Continuum-Source}, and the
disc-wind parameters are summarised in
Table~\ref{tab:std_discwind_parameters}. Although the line is potentially
sensitive to the temperature structure of the wind, determination of a
self-consistent wind temperature is beyond the scope of this
paper. Readers are refereed to \citet{hartmann:1982} in which the wind
temperature structure is determined by balancing the radiative cooling
rate (assuming optically thin) with the MHD wave heating rate.
We pragmatically assume that the wind is isothermal at $T_{\mathrm{wind}}$.

Initially we examine the characteristics of the H$\alpha$ profile as a
function of the wind acceleration parameter $\beta$ and the isothermal
wind temperature $T_{\mathrm{wind}}$.  The mass-loss rate is kept
constant at
$\dot{M}_{\mathrm{wind}}=10^{-8}\,\mathrm{M_{\sun}\,yr^{-1}}$.
Figure~\ref{fig:atlas_discwind_only} shows the model profiles computed
for the wind temperatures between $7000\,\mathrm{K}$ and
$9000\,\mathrm{K}$, and those with $\beta$ between $0.2$ and $2.0$,
computed at inclination $i=55^{\circ}$.
The morphology of the profile exhibited by the model changes from
FU-Ori type (c.f.~\citealt{reipurth:1996}) for the models with small
$\beta$ to the stronger emission types which are more commonly seen in
the observed H$\alpha$ (c.f.~\citealt{reipurth:1996}) as the value of
$\beta$ increases.  
Classical P~Cygni profiles are prominent
in the models with lower values of $T_{\mathrm{wind}}$ and $\beta$
(a colder and faster accelerating wind). As the value of $\beta$
increases, the position of the blue-shifted absorption component moves
toward the line centre. The blue edge of the absorption component in
the fastest accelerating wind 
($\beta=0.2$) model is almost at the terminal velocity ($v_{\infty}\sim
160\,\kmps$) of the disc wind (c.f.~Figure~\ref{fig:discwind_vr_rho}), that
for the slowest wind acceleration model 
($\beta=4.0$) is located close the line centre ($\sim 20\,\kmps$).

 As one can see in Figure~\ref{fig:discwind_vr_rho}, the column density of the
gas moving at $v_{\infty}$ increases as the wind acceleration becomes
larger since the gas moving at $v_{\infty}$ becomes closer to the
photosphere. At the location where the absorption occurs, the gas is
moving faster for the models with the faster wind acceleration models;
hence, the location of the absorption component is bluer as $\beta$
becomes smaller (as the wind acceleration faster). The figure also
shows that the intensity of the emission component increases as
$\beta$ increases for a given $T_{\mathrm{wind}}$. This is mainly
because the density near the photosphere increases as the value of
$\beta$ increases (Figure~\ref{fig:discwind_vr_rho}); hence, the
emissivity of increases. Unlike the magnetospheric accretion only
models (Figure~\ref{fig:atlas_acc_models}), the inverse P~Cygni
profiles are not seen here, nor the blue-ward asymmetry. Because of
the absorption component in P~Cygni profile, most of the profiles show
red-ward asymmetry except for the fast wind acceleration models
($\beta=0.2$ and $0.5$)with low inclinations (not shown here).

    \includegraphics[%
      clip,
      scale=0.45]{figures/discwind_d_effect.eps}

  \caption{The dependency of the H$\alpha$ profile on the source
    displacement parameter $d$. The line profiles are computed with 
    the wind parameter $T_{\mathrm{wind}} = 9000\Kelvin$ and $\beta  =
    0.5$ for $i=55^{\circ}$.  The profiles in the figure uses $d = 
    11~R_{*}$ (dashed),   $18~R_{*}$ (dash-dot) and $22~R_{*}$
    (dash-dot). All the other parameters are same as in
    Figure~\ref{fig:atlas_discwind_only}.  As the wind becomes more
    collimated (larger $d$), the wind launching angle becomes smaller
    and the wind density is enhanced towards the symmetry axis
    ($z$); hence, the wind emission increases.  The absorption
    component is relatively insensitive to $d$ since the change in the
    column density toward the observer is relatively small.  No
    magnetospherical component is included. 
  }

  \label{fig:discwind_d_effect}

    \includegraphics[%
      clip,
      scale=0.45]{figures/discwind_mdot_effect.eps}

  \caption{The effect of the wind mass-loss rate.  The profiles computed
    with $\left(T_{\mathrm{wind}},~\beta \right) = \left( 9000\Kelvin,
    0.5\right)$, and $\dot{M}_{\mathrm{wind}} = 
    2.0 \times 10^{-8}~\mathrm{M_{\sun}\,yr^{-1}}$\, (dashed), 
   $1.0 \times 10^{-8}~\mathrm{M_{\sun}\,yr^{-1}}$\, (solid), and 
   $0.5 \times 10^{-8}~\mathrm{M_{\sun}\,yr^{-1}}$\, (dash-dot) 
    are shown above. The
    models are computed with only the disc-wind component (no magnetospheric
    accretion). With all other parameter fixed
    constant, the mass-loss rate behaves as a scaling factor for the wind
    density (c.f.~equation~\ref{eq:discwind-density}).}

  \label{fig:discwind_mdot_effect}

\end{figure}

The effect of varying the wind mass-loss rate is shown in
Figure~\ref{fig:discwind_mdot_effect}.  The line is very sensitive to the
mass-loss rate. With all other parameters fixed constant,
$\dot{M}_{\mathrm{wind}}$ acts as an scaling factor of the wind
density (equation~\ref{eq:discwind-density}). The important processes
in forming H$\alpha$ emission are the recombination process and the
subsequent downward transition to $n=3$ level. Assuming the ionisation
fraction remains constant, the recombination rate is proportional to
the square of wind density; hence the line is very sensitive to the
change in $\dot{M}_{\mathrm{wind}}$.  We have found the line flux of
the models in the figure is approximately proportional to the square
of the mass-loss rate. This very sensitive nature of H$\alpha$ would
make the line a good candidate for a mass-loss rate indicator;
however, without solving the ambiguity of the emission component from
the magnetospheric accretion flow and the wind, and without knowing
the detailed temperature structure of the wind, it could provide us a
misleading a mass-loss rate estimate of a T Tauri star.

By inspecting the model H$\alpha$ images computed (no shown here), we have
 found that the extent of the H$\alpha$ emission region is relatively
 small i.e.~$h\lesssim 0.2\,\mathrm{au}$ or $20\, R_{*}$ where h is the
 vertical distance from the disc when compared to the
 spectro-astrometric observations of \citet{takami:2003} 
for RU~Lup and CS Cha which show 1--5~au scale outflows%
\footnote{Their observations also show some objects
(e.g~Z CMa and AS~353A) displaying the outflows in larger scale
($>50\,\mathrm{au}$); however, this could be formed in shocks rather
than MHD-wave heating (e.g. \citealt{hartmann:1982}) or X-ray heating
(e.g.~\citealt{shang:2002}). %
}. The extent of the line emission region becomes slightly
larger ($\sim0.4$~au) if a slower wind acceleration rate (e,g,~$\beta=4$)
is used.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}

\caption{A summary of the reference disc-wind model parameters used
in section~\ref{sub:result-discwind-knigge}. See also
section~\ref{sub:model-disc-wind}.}

\begin{tabular}{ccccccc}
\hline 
$d$&
$R_{\mathrm{wi}}$&
$R_{\mathrm{wo}}$&
$\dot{M}_{\mathrm{wind}}$&
$p$&
$f$&
$R_{\mathrm{s}}$\tabularnewline
$\left[R_{*}\right]$&
$\left[R_{*}\right]$&
$\left[\mathrm{au}\right]$&
$\left[M_{\sun}\,\mathrm{yr^{-1}}\right]$&
$[-]$&
$[-]$&
$\left[R_{*}\right]$\tabularnewline
\hline
$22$&
$3.0$&
$1.0$&
$10^{-8}$&
$-3/2$&
$2.0$&
$30$\tabularnewline
\hline
\end{tabular}

\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Disc-wind-magnetosphere hybrid models}
\label{sub:result-discwind-hybrid}

@@

  \caption{The disc-wind-magnetosphere hybrid model. These models
  (solid lines) have the  same parameters as
    Figure~\ref{fig:atlas_discwind_only}, but  also
    include the magnetospheric accretion flow
    ($T_{\mathrm{\max}}=7500~\mathrm{K}$ and
    $\dot{M}_{\mathrm{acc}}=10^{-7}\,\mathrm{M_{\sun}\,yr^{-1}}$). While
    the emission from the magnetosphere dominates for the models with
    smaller $\beta$ and $T_{\mathrm{wind}}$, the emission from the wind dominates the
    profiles for models with larger $\beta$ and $T_{\mathrm{wind}}$.  The contribution
    of the magnetosphere can be judged from the profiles computed with
    the magnetosphere only (dotted lines).}

  \label{fig:atlas_hybrid}

@@

We now consider the models computed with a combination of the
disc-wind and the magnetospheric accretion components. The parameters
used for the magnetosphere are  same as in 
Figure~\ref{fig:acc_inclination_effect}, and  the mass-loss rate of
the disc wind is 
$\dot{M}_{\mathrm{wind}}=10^{-8}\,\mathrm{M_{\sun}\, yr^{-1}}$
(i.e.~$\mu=\dot{M}_{\mathrm{wind}}/\dot{M}_{\mathrm{acc}}=0.1$).
Figure~\ref{fig:atlas_hybrid} shows the models profiles
computed using the same ranges of the disc-wind temperature and the
wind acceleration parameter as in the previous cases.

As in the disc-wind only models
(Figure~\ref{fig:atlas_discwind_only}), the location of the absorption
component moves toward the line centre as the value of $\beta$
increases.  The figure shows that in the models with smaller $\beta$
and $T_{\mathrm{wind}}$, the line emission from the magnetosphere
dominates, while this situation reverses for the models with larger
$\beta$ and $T_{\mathrm{wind}}$. The model with
$\left(T_{\mathrm{wind}},\,\beta\right)=\left(9000\,\mathrm{K},\,2.0\right)$
results in an H$\alpha$ profile that is far too strong to be
compatible with observations (\citealt{reipurth:1996} and
\citealt{alencar:2000}); however, the lower mass-loss rate models
(while keeping $\mu$ constant) produce the profiles with the line
strengths comparable to observations.  Although the models in
Figure~\ref{fig:atlas_hybrid} are computed with limited
ranges of $T_{\mathrm{wind}}$, $\beta$ and a fixed $i$, the resulting
profiles exhibit a wide variety of line profile shapes, many of which
are similar to the types of H$\alpha$ profiles seen in the observations
(c.f.~\citealt{reipurth:1996}).

An examination of the profiles in
Figures~\ref{fig:atlas_discwind_only} and
Figure~\ref{fig:atlas_hybrid} reveals some of problems
with uniqueness that must inevitably occur when trying to assign an
individual profile as wind or accretion dominated i.e. very similar
profiles may arise from very different circumstellar geometries. For
example, the disc-wind only model with
$\left(T_{\mathrm{wind}},\,\beta\right)=\left(8000\,\mathrm{K},\,1.0\right)$
in Figures~\ref{fig:atlas_discwind_only} and the accretion dominated
model with
$\left(T_{\mathrm{wind}},\,\beta\right)=\left(9000\,\mathrm{K},\,0.5\right)$
in Figure~\ref{fig:atlas_hybrid} give very similar
profiles.

@@

    \includegraphics[%
      clip,
      scale=0.45]{figures/discwind_mu_effect.eps}

@@

  \caption{The effect of the mass-loss rate to mass-accretion rate
    ratio ($\mu=\dot{M}_{\mathrm{wind}}/\dot{M}_{\mathrm{acc}}$).  The
    H$\alpha$ profile computed with only the magnetospheric accretion
    (thick solid) is compared with the disc-wind-magnetosphere
    hybrid models for $\mu =$ 0.025 (dash-dot), 0.1
    (solid) and 0.2 (dash). The magnetosphere used here has
    $T_{\mathrm{max}}=7500\,\Kelvin$ and
    $\dot{M}_{\mathrm{acc}}=10^{-7}\,\mathrm{M_{\sun}\,yr^{-1}}$.  The
    temperature and the acceleration parameter of the disc wind are
    $T_{\mathrm{wind}}=9000\,\Kelvin$ and $\beta=0.5$ respectively.
    All the models are computed with the inclination angle
    $i=55^{\circ}$. As $\mu$ increases the emission component becomes
    stronger and the wings becomes broader, as expected. The
    absorption component is relatively insensitive to the value of
    $\mu$ in this example. The wide range of $\mu$ reasonably
    reproduces the profiles similar to ones seen in observations
    (e.g.~\citealt{reipurth:1996}).}

  \label{fig:discwind_mu_effect}

The effect of changing the ratio of the wind mass-loss rate to
mass-accretion rate
($\mu=\dot{M}_{\mathrm{wind}}/\dot{M}_{\mathrm{acc}}$) is demonstrated
in Figure~\ref{fig:discwind_mu_effect}. With a fixed value of
mass-accretion rate
($\dot{M}_{\mathrm{acc}}=10^{-7}\,\mathrm{M_{\sun}\, yr^{-1}}$), the
mass-loss rate is varied. The figure shows that as $\mu$ increases the
P-Cygni absorption deepens only slightly., and the position of minimum flux in the
absorption trough appears to move blue-ward.  On the other hand, the
flux peak increase and the line wings become stronger as $\mu$
increases.  Although MHD models suggests (e.g.~\citealt{koenigl:2000})
that $\mu \sim 0.1$, the figure suggest that models a relatively wide range
of $\mu$  (e.g.~between 0.025 and 0.2 in this case)  reproduce the
profiles similar to observations.

Apparently, the type of profiles  
(affected by the blue-shifted absorption feature) seen in this section
cannot be explained by the magnetospheric accretion flow alone
(e.g.~\citealt{muzerolle:2001}).  Comparison of the model profiles
with different types of the observed H$\alpha$ profiles will be given
in section~\ref{sub:Classification-scheme-proposed}.

@@

\subsection{Inclination dependency of H$\alpha$ profiles}
